Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-19 Manuscript/sec2-review-of-graph-neural-network.tex
更新GAT的公式.

Different GNNs adopt different kinds of graph neurons and have different definitions of the three functions.
$\phi$ and $\Sigma$ are the \emph{edge calculation} functions.
They are conducted over every edge in $\mathcal{G}$.
$\gamma$ is the \emph{vertex calculation} function.
It is conducted over every vertex in $\mathcal{G}$.
\tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} list the edge functions and the vertex functions of typical GNNs, respectively.
For ChebNet, we report its GNN sub-layer in the tables. 
A ChebNet layer consists of $K$ GNN sub-layers and a summation layer:
$\boldsymbol{H}^{l+1} = \sum_{k=1}^K{\boldsymbol{Z}^{(k)} \textcolor{blue}{\boldsymbol{W}^{(k)}}}$ with the GNN sub-layers $\boldsymbol{Z}^{(1)}=\boldsymbol{H}^l$, $\boldsymbol{Z}^{(2)}=\hat{\boldsymbol{L}}\boldsymbol{H}^l$, and $\boldsymbol{Z}^{(k)}=2\hat{\boldsymbol{L}}\boldsymbol{Z}^{(k-1)} - \boldsymbol{Z}^{(k-2)}$, where $\boldsymbol{H}^l$ is the matrix of output hidden feature vectors of the layer $l-1$.
For GAT, a GAT layer consists of two sub-layers:
the first sub-layer calculates the summation $\sum_{k \in \mathcal{N}(i)}\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l,k}_i \parallel \hat{\boldsymbol{h}}^{l,k}_j] ))$, and the second sub-layer calculates $\alpha_{j, i}^k$ and $\boldsymbol{m}^l_{j,i}$.
The vertex calculation of a GAT layer is decomposed into two parts: $\hat{\boldsymbol{h}}^{l,k}_{i}$/$\boldsymbol{h}^{l+1}_i$ is calculated for every vertex before/after the message-passing, respectively.
The edge calculation of GAT uses $\hat{\boldsymbol{h}}^{l,k}_{i}$ during the message-passing.
For GaAN, a GaAN layer consists of four sub-layers: the first sub-layer calculates the summation $\sum_{k \in \mathcal{N}(j)}\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}}  \boldsymbol{h}_k^l] )$, and the other three layers calculate $\boldsymbol{m}^l_{j,i,1}$/$\boldsymbol{m}^l_{j,i,2}$/$\boldsymbol{m}^l_{j,i,3}$.

@@

						 & \alpha_{j, i}^k = \frac {\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l,k}_i \parallel \hat{\boldsymbol{h}}^{l,k}_j] ))} {\sum_{k \in \mathcal{N}(i)}\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l,k}_i \parallel \hat{\boldsymbol{h}}^{l,k}_j] ))} \\
						 & \text{Multi-head concatenation}: \boldsymbol{m}_{j, i}^l = \parallel_{k=1}^K \delta(\alpha_{j, i}^k \hat{\boldsymbol{h}}^{l,k}_j)                                                                                                                                             \\
						 & \text{Multi-head average}: \boldsymbol{m}_{j, i}^l = \frac{1}{K} \sum_{k=1}^K \delta(\alpha_{j, i}^k \hat{\boldsymbol{h}}^{l,k}_j)

@@

			\makecell[r]{concat: $O(d_{in}*d_{out})$ \\
                 average: $O(K * d_{in} * d_{out})$  \\ Two sub-layers} 

@@

			\makecell[l]{$\hat{\boldsymbol{h}}^{l,k}_{i} = \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_i^{l}$ (pre edge calculation),\\ $\boldsymbol{h}_i^{l+1} = \boldsymbol{s}_i^l$}                                                         &
			\makecell[r]{concat: $O(d_{in}*d_{out}/K)$\\ average: $O(d_{in}*d_{out})$
             \\ Two sub-layers}

@@

\textbf{GAT} (low vertex \& high edge calculation complexity):
GAT introduces the attention mechanism into the graph neural networks.
During the message passing, it dynamically determines the weight for an edge based on the hidden feature vectors of its two end vertices.
Moreover, it adopts the multi-head mechanism that generates $K$ independent views for an edge, where $K$ is a hyperparameter.
GAT moves the vertex update function $\gamma$ before the edge calculation.
$\hat{\boldsymbol{h}}^{l,k}_{i} = \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_i^{l}$ with $1 \leq k \leq K$ are conducted for every vertex $v_i$ before the message function $\phi$.
$\phi$ uses $\hat{\boldsymbol{h}}^{l,k}_{i}$ to get the weight of each edge in every head $k$.
After the edge calculation, GAT uses the aggreated vector $\boldsymbol{s}^l_i$ directly as the output hidden vector.

>>>>

Different GNNs adopt different kinds of graph neurons and have different definitions of the three functions.
$\phi$ and $\Sigma$ are the \emph{edge calculation} functions.
They are conducted over every edge in $\mathcal{G}$.
$\gamma$ is the \emph{vertex calculation} function.
It is conducted over every vertex in $\mathcal{G}$.
\tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} list the edge functions and the vertex functions of typical GNNs, respectively.
For ChebNet, we report its GNN sub-layer in the tables. 
A ChebNet layer consists of $K$ GNN sub-layers and a summation layer:
$\boldsymbol{H}^{l+1} = \sum_{k=1}^K{\boldsymbol{Z}^{(k)} \textcolor{blue}{\boldsymbol{W}^{(k)}}}$ with the GNN sub-layers $\boldsymbol{Z}^{(1)}=\boldsymbol{H}^l$, $\boldsymbol{Z}^{(2)}=\hat{\boldsymbol{L}}\boldsymbol{H}^l$, and $\boldsymbol{Z}^{(k)}=2\hat{\boldsymbol{L}}\boldsymbol{Z}^{(k-1)} - \boldsymbol{Z}^{(k-2)}$, where $\boldsymbol{H}^l$ is the matrix of output hidden feature vectors of the layer $l-1$.
For GAT, a GAT layer consists of two sub-layers:
the first sub-layer calculates the summation $\sum_{k \in \mathcal{N}(i)}\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l,k}_i \parallel \hat{\boldsymbol{h}}^{l,k}_j] ))$, and the second sub-layer calculates $\alpha_{j, i}^k$ and $\boldsymbol{m}^l_{j,i}$.
The vertex calculation of a GAT layer is also decomposed into two parts: $\hat{\boldsymbol{h}}^{l,k}_{i}$/$\boldsymbol{h}^{l+1}_i$ is calculated for every vertex before/after the message-passing, respectively.
The edge calculation of GAT uses $\hat{\boldsymbol{h}}^{l,k}_{i}$ during the message-passing.
For GaAN, a GaAN layer consists of four sub-layers: the first sub-layer calculates the summation $\sum_{k \in \mathcal{N}(j)}\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}}  \boldsymbol{h}_k^l] )$, and the other three layers calculate $\boldsymbol{m}^l_{j,i,1}$/$\boldsymbol{m}^l_{j,i,2}$/$\boldsymbol{m}^l_{j,i,3}$.

@@

                         & \text{Sub-layer 0:}\\
                         & \boldsymbol{m}^{l,0}_{j,i,(k)} = \exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l}_{i,(k)} \parallel \hat{\boldsymbol{h}}^{l}_{j, (k)}]))\\
                         & \text{Sub-layer 1:}\\
						 & \alpha_{j, i, (k)} = \frac {\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l}_{i,(k)} \parallel \hat{\boldsymbol{h}}^{l}_{j,(k)}] ))} {\hat{\boldsymbol{a}}^{l,0}_{i,(k)}} \\
						 & \text{Multi-head concatenation}: \boldsymbol{m}_{j, i}^{l,1} = \parallel_{k=1}^K \delta(\alpha_{j, i, (k)} \hat{\boldsymbol{h}}^{l}_{j,(k)})                                                                                                                                             \\
						 & \text{Multi-head average}: \boldsymbol{m}_{j, i}^{l,1} = \frac{1}{K} \sum_{k=1}^K \delta(\alpha_{j, i, (k)} \hat{\boldsymbol{h}}^{l}_{j,(k)})

@@

             $
             \begin{aligned}[t]
                \text{concat: } O(d_{out}) &\\
                \text{average: } O(K * d_{out}) &\\
                \text{Two sub-layers} &
             \end{aligned}
             $ 

@@

                & \text{Before edge calculation:}\\
                & \hat{\boldsymbol{h}}^{l}_{i,(k)} = \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_i^{l}\\
                & \text{Sub-layer 0:}\\
                & \hat{\boldsymbol{a}}^{l,0}_{i,(k)} = \boldsymbol{s}^{l,0}_{i,(k)}\\
                & \text{Sub-layer 1:}\\
                & \boldsymbol{h}_i^{l+1} = \boldsymbol{s}_i^{l,1}
                \end{aligned}
                $
             $
             \begin{aligned}[t]
             \text{concat:} O(d_{in}*d_{out}) &\\
             \text{average:} O(K*d_{in}*d_{out}) &\\
             \text{Two sub-layers}&
            \end{aligned}
             $

@@

\textbf{GAT} (low vertex \& high edge calculation complexity):
GAT introduces the attention mechanism into the graph neural networks.
Each GAT layer consists of two sub-layers.
The first sub-layer calculates the attention weights of every edge $(v_i, v_j)$ $\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l,k}_i \parallel \hat{\boldsymbol{h}}^{l,k}_j]$ and aggregates the attention weights of all adjacency edges of every vertex $v_i$ as $\sum_{k \in \mathcal{N}(i)}\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l,k}_i \parallel \hat{\boldsymbol{h}}^{l,k}_j] ))$.
The second sub-layer uses 
During the message passing, it dynamically determines the weight for an edge based on the hidden feature vectors of its two end vertices.
Moreover, it adopts the multi-head mechanism that generates $K$ independent views for an edge, where $K$ is a hyperparameter.
GAT moves the vertex update function $\gamma$ before the edge calculation.
$\hat{\boldsymbol{h}}^{l,k}_{i} = \textcolor{blue}{\boldsymbol{W}^{l,k}} \boldsymbol{h}_i^{l}$ with $1 \leq k \leq K$ are conducted for every vertex $v_i$ before the message function $\phi$.
$\phi$ uses $\hat{\boldsymbol{h}}^{l,k}_{i}$ to get the weight of each edge in every head $k$.
After the edge calculation, GAT uses the aggreated vector $\boldsymbol{s}^l_i$ directly as the output hidden vector.

### 2020-07-19 Manuscript/sec4-experiments.tex
更新GAT的公式.

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_gcn.png}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_ggnn.png}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_gat.png}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_gaan.png}}
    \caption{Vertex/edge calculation time-consuming ratio.}
    \label{fig:exp_vertex_edge_cal_proportion}
\end{figure}

>>>>

\begin{figure}
    \centering
    \subfloat[GCN]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_gat.pdf}}
    \subfloat[GaAN]{\includegraphics[height=4cm]{figs/experiments/exp_vertex_edge_cal_proportion_gaan.pdf}}
    \caption{Vertex/edge calculation time-consuming ratio.}
    \label{fig:exp_vertex_edge_cal_proportion}
\end{figure}
